---
title: "Module 1"
author: "Ananya Sharma"
date:   "15/01/2022"
output:
   
     html_document:
           toc: TRUE
           toc_float: TRUE
           theme: cerulean
  
---

```{r setup, include=FALSE, cache=TRUE,error=TRUE}
 options(repos = list(CRAN="http://cran.rstudio.com/"))
 knitr::opts_chunk$set(echo = TRUE)
 
#install.packages("ggplot2")
library(ggplot2)
#install.packages("magrittr")
library(magrittr)
#install.packages("dplyr")
library(dplyr)
#install.packages("psych")
#install.packages("psych", repo="http://cran.r-project.org", dep=T)
library(psych)
#install.packages("tidyverse")
library(tidyverse)
#install.packages("reshape")
library(reshape)
#install.packages('broom')
library(broom)
#install.packages('fastDummies')
library(fastDummies)
#install.packages("ggcorrplot")
library(ggcorrplot)
#install.packages("tidyr")
# install.packages("broom")
 library(broom)
```
 

## Ames Iowa Data Set- Introduction

* This data set contains data on 2931 real estate properties around Ames Iowa and 82    columns describing the property. 
* There are many qualities of a real estate that determine its net worth.    
* In this data set, we are exploring correlation and the parameters that helps people   to understand which variables are the most suitable factors while purchasing a
  house.
* The data set has a lot of nominal and ordinal variables which can be converted to
  meaningful visualizations.
* We are performing correlation analysis and Descriptive analytics in this Module.

## Reading the data set:

```{r }
 getwd()   # Understanding where the file has been saved in the system
 Housing_dataset<-read.csv("AmesHousing.csv")

#Convert to dataframe

Ames_House<-as.data.frame(Housing_dataset)       #Reading the dataset into a variable

```
## Describing the Data Set

* Using the Summary function, we are able to understand that mean Sale Price of Iowa
  House is `$1,80,000` tentatively, whereas the median price of the house is
 `$1,60,000`.
* The dataset has lot of values that are not available, has been left blank or filled
  with zero's and one's in plenty.

## Using Descriptive Statistics for Analysis
```{r,comment=NA}

psych::describe(Ames_House)

```

# Exploratory Data Analysis

## Histogram

* The data is right skewed.The Histogram indicates that the houses that are cheap are   sold more often as compared to the houses that have more price.

* The sale price feature has been log performed to avoid using negative Sales Price.  
  The median Sales price of houses is around 160000. It is also shown by the line  
  going across through the histogram.

```{r,comment=NA}
 library(magrittr)
 library(ggplot2)
 Histogram_House<-Ames_House %>% ggplot(aes(x=SalePrice,fill=Sale.Condition)) + 
 geom_histogram() +
 geom_density(alpha=.2, fill="White") + geom_vline(aes(xintercept=mean(SalePrice)),
            color="black", linetype="dashed", size=1)+ ylab("Frequency")+ xlab("Price of Houses")
 Histogram_House
```


#Boxplot of Sales Price

* Box plot on the sales Price shows that there are a lot of outliers in Sales Vale of    the data set and it has to be cleaned.

```{r,comment=NA}
    Ames_House %>% ggplot(aes(x=SalePrice)) +
     geom_boxplot(fill='pink')+
     theme( legend.position = "none" ) + theme_light()
```

## Creating a Scatter Plot to understand which Neighbourhood has good land value

* We are able to understand from the graph that Noridght and Names have Expensive       Property Prices.
* Old Town has Less expensive property prices.
* Names has around 400 properties whereas Grn hill has 2 properties based out of that
  region.

```{r,comment=NA}

table(Ames_House$Neighborhood)
Ames_House %>% ggplot( aes(x = SalePrice,y=Neighborhood, color=Neighborhood)) + 
geom_point(alpha=0.2) + xlab("Sales Price of House") + ylab("Neighbourhoods") +
ggtitle("Price Vs Neighbourhoods") + scale_x_log10()+ theme_light()

```

## Boxplot- Sale Price vs Ms Zoning & House Style

* The Box plot does a Bi-Variate Analysis where we plot the value of Sale Price vs the   zone areas.
* The areas are zoned as Residential, Agricultural, Industrial, Commercial and   
  Village.
* The graph shows that Village Property is more expensive as compared to Commercial
  Property.
* Agricultural Property has the least property value. 
* The second box plot shows that the unfinished units have the lowest value.
* All the units are priced between `$1,00,000` to `$3,00,000`.

```{r,comment=NA}
table(Ames_House$MS.Zoning)   #Using the table function to Understand the count of categorical data

  ggplot(Ames_House, aes(x = MS.Zoning, y = SalePrice, fill = MS.Zoning)) + 
  geom_boxplot(alpha=0.4) + theme(legend.position="top")+ ggtitle("Zone Vs Sale Price") +
  stat_summary(fun = "mean", geom = "point", shape = 6,size = 1)
  
  table(Ames_House$House.Style)  # Understand the count of House Styles

  ggplot(Ames_House, aes(x = House.Style, y = SalePrice, fill = House.Style)) + 
  geom_boxplot(alpha=0.3) + theme(legend.position="right")+ ggtitle("House Style vs Sale Price") +
  stat_summary(fun = "mean", geom = "point", shape = 1,size = 2) 

```

# Stacked Bar Plot. 

* Major of the real estate properties have Good Basement Quality and Typical
  Structure.

```{r, echo=FALSE,comment=NA}
Ames_House %>% ggplot( aes(x = Bsmt.Cond, fill = Bsmt.Qual),position = position_dodge(preserve = 'single')) +geom_bar()+ xlab("Basement Condition") +ylab("Frequency")
```

##Analyzing Correlation

* From the subset of Numerical and Continuous values taken from the Ames Housing data
  set we find      that Over all Quality of the House(0.799) and Gr Living area(0.706)
  have a Strong Correlation with Sale Price.
* Sale Price and Fireplace have a weak correlation with 0.558 degrees of relationship 
* Sale Price and Screen Porch has a super weak relationship(0.11)

```{r,echo=FALSE, comment=NA}
library(ggcorrplot)
library(dplyr)
library(tidyverse)
library(tidyr)

#Creating Subset for analyzing correlation amongst the variables
colnames(Ames_House)    #To check which variables we can take for corrplot
#Considering a few numerical values 
c<-subset(Ames_House,select=c(SalePrice,Year.Built,Screen.Porch,Overall.Qual,Lot.Area,Fireplaces,Gr.Liv.Area))
MyCorr1<-as.data.frame(cor(c))  #Converting to a data frame
                 
cor(MyCorr1)
ggcorrplot(cor(c),method = "circle",
                       title = "Correlation Plot:Table for Numerical Data"
                       ,type = "upper",
                       lab = TRUE)

y<-rbind(cbind(Description = "Lot Area and sale price",
               tidy(cor.test(Ames_House$SalePrice,Ames_House$Lot.Area))),
               cbind(Description = "Overall Quality and Sale Price",
               tidy(cor.test(Ames_House$SalePrice, Ames_House$Overall.Qual
                             ))))  %>%
  dplyr::select(Description,R = estimate, p.value, statistic , conf.low, conf.high) %>%
  mutate(
    R = round(R,3),
    statistic = round(statistic,3),
    conf.low = round(conf.low,3),
    conf.high = round(conf.high,3)
  )

print(y)

```

## Plotting Heat Map

* The categorical value have to be converted to factors or dummy variables with 
  nominal values before  plotting graph or correlation. However we must avoid
  categorical values for getting the correlation.
* Not a very clear plot, however, it sets the pace for further analysis.

```{r,comment=NA}
# Coverting the ordinal variables to factors
Ames_House$Exter.Cond <-  as.numeric(factor(Ames_House$Exter.Cond, 
                                  levels = c("Ex", "Fa","Gd", "TA","Po"),
                                  labels = c(5,2,4,3,1) ,ordered = TRUE))

Ames_House$Heating.QC <-  as.numeric(factor(Ames_House$Heating.QC, 
                                  levels = c("Ex", "Fa","Gd", "TA","Po"),
                                  labels = c(5,2,4,3,1) ,ordered = TRUE))
Ames_House$Central.Air <- as.numeric(factor(Ames_House$Central.Air, 
                                  levels = c("N", "Y"),
                                  labels = c(0,1) ,ordered = TRUE))

model_var <- c('SalePrice', 
                'Overall.Qual','Exter.Cond','Overall.Cond','Total.Bsmt.SF','Heating.QC', 
                'Central.Air','Gr.Liv.Area','Bedroom.AbvGr','Kitchen.AbvGr',
                'TotRms.AbvGrd','Garage.Area','Open.Porch.SF','Yr.Sold')

heat <- Ames_House[,model_var]

options(repr.plot.width=8, repr.plot.height=6)

qplot(x=X1, y=X2, data=melt(cor(heat, use="p")), fill=value, geom="tile") +
   scale_fill_gradient2(low = "Blue", high = "pink", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Correlation") +
   theme_minimal()+ 
   theme(axis.text.x = element_text(angle = 45, vjust = 1, size = 8, hjust = 1))+
   coord_fixed()+
   ggtitle("Figure 7 Correlation Heatmap") +
   theme(plot.title = element_text(hjust = 0.3))
```
## Heat Map for Numerical Dataset

* This heat map shows the correlation of all numerical variables.
* This is not a very clean map but it sets the overall sentiment as to which variable
  could be strongly related and which one would have weak correlation via a graphical
  representation.
  
```{r,comment=NA}
#using the plotly function to plot the graph

library(plotly)
library(reshape2)
  AmesData_Numeric_Vals=Ames_House%>%
  dplyr::select_if(is.numeric)%>%   #converting values to numeric
  drop_na()
  Analyse_Data=cor(AmesData_Numeric_Vals)             
  Analyse_melt=melt(Analyse_Data)     #Converting the value from categorical to numerical
  Correlation_Ames=ggplot(Analyse_melt,mapping=aes(x=X1,y=X2,fill=value))+
  geom_tile()+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(text = element_text(size=9))+
  ggtitle("Heat Map for Ames Iowa Housing Data")+
  ylab("All the factors that are numeric in nature")+
  xlab("All the factors that are numeric in nature")+
  scale_fill_distiller(palette = "BuPu")
  ggplotly(Correlation_Ames, tooltip="text")

```
## Analysis:  

* The data contains 78 prediction variables that pertain to various aspects of an Ames
  home.Continuous, discrete, nominal, and ordinal types are some of these attributes.
  There are 44 category variables, with 21 nominal and 23 ordinal elements.

* The Sales Price has a very strong correlation with Overall Quality of the house and
  Fireplaces has a weak correlation.    

## References:

* ggplot2 histogram plot : Quick start guide - R software and data visualization - 
  Easy Guides - Wiki - STHDA. (2018).
  http://www.sthda.com/english/wiki/ggplot2-histogram-plot-quick-start-guide-r-softwa
  re-and-data-visualization  

* Coder, R. (2021, November 14). Heat map in ggplot2. R CHARTS | A Collection of
  Charts and Graphs Made with the R Programming Language. 
  https://r-charts.com/correlation/heat-map-ggplot2/

* GeeksforGeeks. (2021, December 15). Box plot in R using ggplot2.
  https://www.geeksforgeeks.org/box-plot-in-r-using-ggplot2/
